---
title: "ROBOMICS 16S Analysis"
author: "Wesley Sparagon"
date: "7/9/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries and custom functions.

```{r}
#Load libraries
library(vegan)
library(ggplot2)
library(pairwiseAdonis)

#Load functions
source(file="cull.otu.script.R")
source(file="generate.long.format.script.R")
source(file="generate.square.dist.script.07.27.2020.R")
source(file="zscore.calculation.script.R")
```

Load in data and metadata

```{r}
#load unifrac data
unifrac=read.csv(file="otu_repr_100.tre1.wsummary.csv")

#load relabund
relabund=read.csv(file="abundance_table_100.database.csv")

#load metadat
metadata=read.csv(file="metadata.csv")
```

Work up metadata.

```{r}
#change inocculum and media factor levels
metadata$Innoculum=factor(metadata$Innoculum,levels=c("Water Innoculum", "Coral Innoculum", "Algal Innoculum", "BACKREEF", "Unknown", "None"))
metadata$Media=factor(metadata$Media,levels=c("Water DOM", "Coral DOM", "Algal DOM", "Blank", "None"))
metadata$Timepoint..hours.=as.factor(metadata$Timepoint..hours.)

#subset for just sterivex samples
metadata1=metadata[metadata$Sample_Type=="Sterivex",]

#subset metadata1 for just Aquaria and Backreef type samples
metadata.inoc=subset(metadata1, Item=="Aquaria" | Item=="Backreef")

#subset metadata.ino for just t0 samples.
metadata.inoc.t0=metadata.inoc[metadata.inoc$Timepoint..hours.=="0",]

#subset metadata1 for t0 innoculum and t0 bottle samples.
metadata.t0=metadata1[metadata1$Timepoint..hours.=="0",]
metadata.t0.v1=metadata.t0[-37,] #remove unknown sample

```

Work up unifrac.

```{r}
#generate squre dist
generate.square.dist.matrix(unifrac,metadata.inoc,1)
unifrac.matrix.inoc=dist.matrix1 #rename output
unifrac.dist.inoc=square.dist.matrix #rename output

#generate squre dist for inoc t0
generate.square.dist.matrix(unifrac,metadata.inoc.t0,1)
unifrac.matrix.inoc.t0=dist.matrix1 #rename output
unifrac.dist.inoc.t0=square.dist.matrix #rename output

#generate unifrac for t0 samples
generate.square.dist.matrix(unifrac,metadata.t0.v1,1)
unifrac.matrix.t0=dist.matrix1 #rename output
unifrac.dist.t0=square.dist.matrix #rename output

```

Generate NMDS.

```{r}
nmds.inoc=metaMDS(unifrac.dist.inoc,k=2,trymax=100)
```

Work up NMDS.

```{r}
#extract scores
nmds.inoc.scores=as.data.frame(scores(nmds.inoc))

#add samplename column
nmds.inoc.scores$SampleName=rownames(nmds.inoc.scores)

#merge in metadata
nmds.inoc.scores1=merge(nmds.inoc.scores,metadata.inoc,by.x="SampleName",by.y="Sample_ID")
```

Plot nmds in ggplot2.

```{r}
ggplot(nmds.inoc.scores1,aes(x=NMDS1,y=NMDS2,color=Innoculum,shape=Timepoint..hours.))+
  geom_point(size=6)+
  geom_line(aes(group=Innoculum_Replicate,color=Innoculum))
  
```

NMDS looks good! Innoculum changes/differentiates from t-48 to t0, seperates by innoculum type, and water innoculum is most similar to the backreef samples (and differentiated the least from t-48 to t0).

Next, test the effect of innoculum on t0 aquaria.

```{r}
permanova.inoc.t0 = adonis2(unifrac.dist.inoc.t0 ~ Innoculum, data=metadata.inoc.t0, permutations=999)
permanova.inoc.t0 #innoculum is significant!

#test pairwise comparisons
pairwise.permanova.inoc.t0=pairwise.adonis(unifrac.dist.inoc.t0, factors = metadata.inoc.t0$Innoculum, p.adjust.m = "BH")
pairwise.permanova.inoc.t0
```

Bacterial communities differ significantly by innoculum (as expected), although there were no significant pairwise differences, likely due to low sample size.

Next, remove rare taxa.

```{r}
#work up relabund df.
relabund1=relabund[,-234:-236] #remove taxonomy columns
rownames(relabund1)=relabund1$OTUNumber #update rownames
relabund2=relabund1[,-1] #remove otu names column
relabund2.t=as.data.frame(t(relabund2)) #transpose relabund2

#cull otu
cull.otu(relabund2.t,3,.005,.01) #451 ASVs remain.

#visualize ASV idstibution.
hist(unlist(relabund2.t)) #with zeros
hist(unlist(relabund2.t)[unlist(relabund2.t)!=0]) #without zeros
hist(unlist(relabund.df.cull)[unlist(relabund.df.cull)!=0]) #after culling

#save relabund.df.cull as new name.
relabund2.t.cull=relabund.df.cull
```


Next, plot a stacked barchart of the innoculum and t0 bottle.

```{r}
#First, prep data frames for long format script.

#work up relabund2.t.cull
relabund2.cull=as.data.frame(t(relabund2.t.cull)) #transpose

#work up taxonomy df
taxonomy=relabund[,234:236] #extract taxonomy coumns.
taxonomy1=cbind(taxonomy,t(as.data.frame(strsplit(as.character(taxonomy$OTUConTaxonomy),split=";")))) #split contaxonomy by ";", add to taxonomy df
rownames(taxonomy1)=taxonomy1$repSeqName #change rownames
colnames(taxonomy1)[c(1,4:9)]=c("OTUNumber","Domain","Phylum","Class","Order","Family","Genus") #update colnames
taxonomy1.cull=taxonomy1[taxonomy1$OTUNumber %in% rownames(relabund2.cull),] #subset for only culled ASVs


generate.long.format(relabund2.cull,metadata.t0.v1,taxonomy1.cull)

#export as csv, generate stacked bar charts in JMP
write.csv(abund.longformat,file="abund.longformat.csv")

```

Next, plot the t0 bottle samples with the innoculum samples. First, generate and work up NMDS.

```{r}
nmds.t0=metaMDS(unifrac.dist.t0,k=2,trymax=100)

#extract scores
nmds.t0.scores=as.data.frame(scores(nmds.t0))

#add sample name column
nmds.t0.scores$Sample_ID=rownames(nmds.t0.scores)

#merge in metadata
nmds.t0.scores1=merge(nmds.t0.scores,metadata.t0.v1)

```

Visualize.

```{r}
ggplot(nmds.t0.scores1,aes(x=NMDS1,y=NMDS2,shape=Item,fill=Innoculum,color=Media))+
  geom_point(size=4)+
  scale_shape_manual(values=c(24,21,0,0,0))+
  scale_fill_manual(values=c("steelblue3","indianred1","springgreen3","black","black","black"))+
  scale_color_manual(values=c("steelblue3","indianred1","springgreen3","black","black"))+
  geom_text(aes(label=Sample_ID))
  
```

Generally looks good. C3 innoculum aquaria and associated bottles appear to be outliers (check lab nb), this is consistent with the previous run. However, C3_w_t00_v1 appaeared to cluster with the rest of the C bottles and not the C3 outlier group. Include C3_w_t00_v1 and remove C3_w_t00.





